---
title: "3rd assignment"
author: "Nischal Pokhrel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
We're going to simulate some data that can be used in each of the following tests.

The response variable is normally distributed, and then we're going to back calculate some predictor variables.

NOTE: There are 9 questions below to answer. Do your best to answer them in full sentences. This is part of what we're practicing in this class.

QUESTION: Describe the trait that you're simulating. This can be a real trait that you are working on, or hope to work on, or it can be completely made up. 
Pay special attention to the N that you're expecting, the mean, and the sd around the mean. 
Answer-I will be working on protein expression in *Staphylococcus aureus*. I want to measure what proportion of celss are involved in protein secretion of vraT protein, the vraT protein is alleged to aid in vancomycin resistance.

```{r simulate vraT expression level}
rnorm(1000, 50, 1)->vraT_expression_level 

plot(density(vraT_expression_level))
min(vraT_expression_level)
max(vraT_expression_level)

```
