---
title: "3rd assignment"
author: "Nischal Pokhrel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
We're going to simulate some data that can be used in each of the following tests.

The response variable is normally distributed, and then we're going to back calculate some predictor variables.

NOTE: There are 9 questions below to answer. Do your best to answer them in full sentences. This is part of what we're practicing in this class.

QUESTION: Describe the trait that you're simulating. This can be a real trait that you are working on, or hope to work on, or it can be completely made up. 
Pay special attention to the N that you're expecting, the mean, and the sd around the mean. 

Answer-I will be working on protein expression in *Staphylococcus aureus*. I want to measure what proportion of cells (in a total population of 100 million) are involved in protein secretion of vraT protein, the vraT protein is alleged to aid in vancomycin resistance. In this context, 1 represents 1 million cells, 2 represents 2 million cells, and so on.

```{r simulate vraT expression level}
rnorm(1000, 50, 1)->vraT_expression_level 

plot(density(vraT_expression_level))
min(vraT_expression_level)
max(vraT_expression_level)

```{r simulate predictor variable for a t test}

yes<-sample(vraT_expression_level, 500, replace=FALSE, prob = ifelse(vraT_expression_level>0, 1, 0)) #### play with this line! Is the test statistically significant. When is it not?
predictor_t.test<-(vraT_expression_level %in% yes)

cbind.data.frame(vraT_expression_level, predictor_t.test)->data

mean(data[which(data$predictor_t.test==TRUE),1])
mean(data[which(data$predictor_t.test==FALSE),1])
t.test(vraT_expression_level~predictor_t.test, data=data) ### this does a two sample t-test. What would a one sample t test be testing? How would you do that? Answer-A two sample test would be comparing two mutant strains on *S. aureus*. One sample t-test would compare any mutant strain with the wild type strain.

### plots our two samples for distribution
plot(density(data[which(data$predictor_t.test==FALSE),1]), col="red", main="Two sample t test")
lines(density(data[which(data$predictor_t.test==TRUE),1]), ylim=c(0, 0.1), xlim=c(-20,20), main="Two Sample T test")

###plot one sample distribution
plot(density(data$vraT_expression_level), col="red", main="One sample t test")
t.test(data$vraT_expression_level) ### what is this test doing? This test is doing a 1 sample T-test 


```
```{r Anova}

# Sample groups from vraT_expression_level
strains1 <- sample(vraT_expression_level, 200, replace=FALSE, prob = ifelse(vraT_expression_level > 3, 1, 0))
strains2 <- sample(vraT_expression_level[which(vraT_expression_level %in% strains1 == FALSE)], 200, replace=FALSE, prob = ifelse(vraT_expression_level[which(vraT_expression_level %in% strains1 == FALSE)] > 4, 1, 0))
strains3 <- sample(vraT_expression_level[which(vraT_expression_level %in% strains2 == FALSE | vraT_expression_level %in% strains1 == FALSE)], 200, replace=FALSE, prob = ifelse(vraT_expression_level[which(vraT_expression_level %in% strains2 == FALSE | vraT_expression_level %in% strains1 == FALSE)] > 5, 1, 0))
strains4 <- sample(vraT_expression_level[which(vraT_expression_level %in% strains2 == FALSE | vraT_expression_level %in% strains1 == FALSE | vraT_expression_level %in% strains3 == FALSE)], 200, replace=FALSE, prob = ifelse(vraT_expression_level[which(vraT_expression_level %in% strains2 == FALSE | vraT_expression_level %in% strains1 == FALSE | vraT_expression_level %in% strains3 == FALSE)] > 6, 1, 0))
strains5 <- vraT_expression_level[which(vraT_expression_level %in% strains1 == FALSE | vraT_expression_level %in% strains2 == FALSE | vraT_expression_level %in% strains3 == FALSE | vraT_expression_level %in% strains4 == FALSE)]

# Plot density distributions for each strains group
plot(density(strains1), ylim=c(0, 1), main="Anovas")
lines(density(strains2), col="red")
lines(density(strains3), col="blue")
lines(density(strains4), col="purple")
lines(density(strains5), col="yellow")

# Create a new data frame for ANOVA groups
anova_predictor <- data.frame(ifelse(vraT_expression_level %in% strains1 == TRUE, "group1", 
                                      ifelse(vraT_expression_level %in% strains2 == TRUE, "group2", 
                                      ifelse(vraT_expression_level %in% strains3 == TRUE, "group3", 
                                      ifelse(vraT_expression_level %in% strains4 == TRUE, "group4", 
                                      "group5")))))

# Combine data into one data frame
data2 <- cbind.data.frame(data, anova_predictor)
names(data2) <- c("vraT_expression_level", "predictor_t.test", "anova_predictor")

# Perform ANOVA analysis
anova(aov(vraT_expression_level ~ anova_predictor, data=data2))  ### what does this do? Answer-It performs analysis of variance

anova(lm(vraT_expression_level ~ anova_predictor, data=data2))    ### what does this do? Answer-It fits a linear model before doing ANOVA
summary(aov(vraT_expression_level ~ anova_predictor, data=data2))  ### what does this do? What do you notice about the last three tests? It provides a summary of the results. All three test are generating F and p-values which also shows that these samples don't show significant difference.
### what information is missing here that you wished you had to understand your study better?
analysis_of_variance<-aov(vraT_expression_level~anova_predictor, data=data2) ##name the model to keep it for downstream
TukeyHSD(analysis_of_variance, conf.level = 0.95) ### what does this do, and where are the differences? If the ANOVA test led to rejection of the null hypothesis than we performed this test to see which group is the one that is different than the rest. 
```
QUESTION: Write one sentence where you report your ANOVA and Tukey tests. What did you find, and how do you report this?
Answer-The analysis of variance showed that there is no difference between each groups and despite this I performed a Tukey test which aligns with the ANOVA.


Again, our simulations aren't to be actually causal, I'm simulating predictor variables to fit a response I've already made. Normally, we would simulate the response variables from more thoughtful predictor variables.

QUESTION:what is the difference between the assumed distributions for the prior predictor variables, and this one?
Answer-The prior predictor variables were based on normal distribution but these are probably biased.

QUESTION: Describe the predictor variable. What does this mean biologically to your trait of interest. How did you decide on the numbers in line 104?

Answer- The predictor variable is the expression level of vraT protein in different strains of *S. aureus*

QUESTION: What is the difference between a regression and a correlation? When would you use each? How does the test stat from the correlation compare to the effect size from the regression? Answer-Regression are equations we can use to predict response variable and correlation gives us how much of effect can a predictor variable have on response variable. I would use regression to predict a response variable without extrapolation from the regression line and I would use correlation to define the overall effect of response variable on predictor variable. 

```{r Linear Regression and correlation}

mutation_density<-1*vraT_expression_level+rnorm(1000, 0, 4) ### change these numbers!! Remember that this is backwards from how we did this on day 1, so the slope should go the other way!
data3<-cbind(data2, mutation_density)

lm<-lm(vraT_expression_level~mutation_density, data=data3)
summary(lm)### what is the output here? What are we interested in understanding from this model? How do we get m? How do we get the intercept?

eq = paste0("y = ", round(lm$coefficients[2],1), "*x", "+",round(lm$coefficients[1],1), ", R^2=", round(summary(lm)$adj.r.squared, digits=3))
plot(data3$mutation_density, data3$vraT_expression_level, col="red", main=eq)
abline(lm, col="black")

### with the same data:
cor.test(data3$vraT_expression_level, data3$mutation_density) ### compare the sample estimate to the 1) simulated effect sizes and 2) to the estimated effect size 
### how does the correlation estimate change when you change line 104?
```

QUESTION: Report your regression and correlation in a sentence. Differentiate between them and what you report for each. 
Answer- The correlation coefficient is positive although the R^2 shows that only a little variance is explained by the predictor variable.
